Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{{DISTRIBUTION STATEMENT A. Approved for public release.} }

### 2021-05-14 paper/synthesis.tex
adding contig type info for filter

Synthesis maps from model and specifications to code. The synthesis algorithm traverses the system architecture looking
for occurrences of filter and monitor specifications. For each such 
occurrence it generates a CakeML program. In the following, we examine both filter and monitor synthesis. The latter is typically much more involved, and we will devote more attention to it.

@@

A filter is intended to be very simple; it is expected to have one
input port and one output; messages on the input that the filter
policy admits pass unchanged to the output port; all others are not
passed on. We have explored in our work two kinds of filter. In the
first, a relatively shallow scan of the input buffer can enforce the
policy. For example, we have used the expressive power of Contiguity
Types \cite{contiguity-types} to enforce \emph{lightweight} bounds constraints
on GPS coordinates in UxAS messages. On the other hand, the second
kind of filter will parse the input buffer into a data structure
specified in AGREE and apply an user-defined \emph{well-formedness}
property, also specified in AGREE, to the data. This allows
arbitrarily complex well-formedness checking.

@@

\emph{Need some discussion of filters and their step-wise propertie s in relation to their infinitary properties. Reference to Johannes' work.}

@@

The synthesized CakeML code for the filter is shown in \figref{fig:filter-cakeml}. The code is called at dispatch by the scheduler. The \texttt{API.callFFI} is the link to the communication fabric to capture input and provide output. The body of the function restates the filter contract to make the appropriate assignments in a way that matches the truth value of the predicate in the filter guarantee.
The auto-generated AGREE specification raises an alert output when the relation is violated.

@@

\paragraph{Initialization}
A monitor may need to accumulate a certain minimum number of observations before being able to make a meaningful assessment of behavior. Until that threshold is attained, the monitor is essentially in a kind of \emph{initialization} phase. In order for correct code to be generated, monitor specifications need to spell out the values of output ports when in such initialization phases. For example, suppose a monitor does some kind of differential assessment of inputs at adjacent timeslices, alerting when (say) the measured location of a UAV at times $t$ and $t+1$ is such that the distance between the two locations is unusually large. Such a monitor needs two measurements before making its first judgement, but at the time of its first output, only one measurement will have been made. The specification has to explicitly say what the correct first output is.

>>>>

Synthesis maps from model and specifications to code. The synthesis
algorithm traverses the system architecture looking for occurrences of
filter and monitor specifications;  for each such occurrence it
generates a CakeML program. In the following, we examine both filter
and monitor synthesis. The latter is typically much more involved, and
we will therefore devote more attention to it.

@@

A filter is intended to be very simple; it is expected to have one
input port and one output; messages on the input that the filter
policy admits pass unchanged to the output port; all others are
dropped (not passed on). We have explored in our work two kinds of
filter. In the first, a relatively shallow scan of the input buffer
can enforce the policy. For example, we have used the expressive power
of Contiguity Types \cite{contiguity-types} to enforce
\emph{lightweight} bounds constraints on GPS coordinates in UxAS
messages. On the other hand, the second kind of filter will parse the
input buffer into a data structure specified in AGREE and apply an
user-defined \emph{well-formedness} property, also specified in AGREE,
to the data. This allows arbitrarily complex well-formedness checking.

@@

{\emph{Need some discussion of filters and their step-wise properties
  in relation to their infinitary properties. Reference to Johannes'
  work.}

\newsavebox{\contig}
\begin{lrbox}{\contig}
\begin{lstlisting}[style=myML]
  Waypoint =
    {Latitude  : f64
     Longitude : f64
     Altitude  : f32
     Check     : Assert
      (~90.0 <= Latitude and Latitude <= 90.0 and
       ~180.0 <= Longitude and Longitude <= 180.0 and
       1000.0 <= Altitude and Altitude <= 15000.0)}

  AutomationResponse =
    {TaskID : i64
     Length : u8
     Waypoints : Waypoint [3]}
\end{lstlisting}
\end{lrbox}

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \scalebox{0.60}{\usebox{\contig}}
    \end{tabular}
  \end{center}
  \caption{Contiguity type specification for filter.}
  \label{fig:filter-contig}
\end{figure}

@@

The contiguity type specification for the filter is shown in
\figref{fig:filter-contig}. The synthesized CakeML code for the filter is shown in
\figref{fig:filter-cakeml}. The code is called at dispatch by the
scheduler. The \texttt{API.callFFI} is the link to the communication
fabric to capture input and provide output. The body of the function
restates the filter contract to make the appropriate assignments in a
way that matches the truth value of the predicate in the filter
guarantee.  The auto-generated AGREE specification raises an alert
output when the relation is violated.

@@

\paragraph{Initialization}
A monitor may need to accumulate a certain minimum number of
observations before being able to make a meaningful assessment of
behavior. Until that threshold is attained, the monitor is essentially
in a kind of \emph{initialization} phase. In order for correct code to
be generated, monitor specifications need to spell out the values of
output ports when in an initialization phase. For example, suppose a
monitor does some kind of differential assessment of inputs at
adjacent timeslices, alerting when (say) the measured location of a
UAV at times $t$ and $t+1$ is such that the distance between the two
locations is unusually large. Such a monitor needs two measurements
before making its first judgement, but at the time of its first
output, only one measurement will have been made. The specification
for must explicitly state what the correct first output is.
